HealpixNet: Efficient spherical Convolutional Neural Network with Healpix sampling for cosmological applications

### 2018-07-08 scnn.tex
update todolist
some edits of healpix sampling

    \item (TK) update the README for data zip - file from NP
    \item Finalise readme on Github code
    \item TK check part sphere notebook and whole sphere
    \item branch with notebooks with plots (results) NP
    \item Michael section 4.5 (efficient convolutions) and 3.2 (conv on graphs)
    \item NP move the repo to HealpixNet

    \item (TK) make the bibliography display only a few name
    % \item (yes) Michael, do you agree with the title HealpixNet? I (Michaël) extensively modified the abstract to reflect that it is a special case of graph CNNs tailored to Healpix. Hence I now agree with the title. Feel free to edit. Ideally I'd like to have "graph" in the title (graph CNNs are quite hot these days), but I couldn't find how.
    % \item Emphasize the novelty of using it on Healpix in abstract, into and rest of the paper
    \item (TK) add info on comparison with histograms, so that it's just a demonstration
    % \item Finish section on Convolutions on graphs (Michael - Nathanael)
    % \item finish section on Efficient convolutions (Michael)
    % \item (done) Plot of the new filters from Mass maps
    % \item (done!, plot to be added) check the power spectrum after smoothing
    % \item (done!) upgrade the plot of pooling (Nathanael)
    \item (TK) possibly improve plot of pooling
	% \item (done!) new plot for network structure (teaser plot) (Michael) What do you think? Todo: link the points on the sphere to the fully connected \nati{I like the plot, but I am thinking about putting multiple spheres with different maps for the convolutiion instead…}
    \item (TK NP) if we have time make a new candidate for network plot (optional)
    % \item (done!) Figure 1 to be round
    \item (first version done) conclusions
    % \item (done!) new simulations with 3 arcmin smoothing
    % \item (done!) simulation with cross-validation for the histogram
    % \item (done!) Update result section according to the new simulations...
    % \item (done!) simulation PSD
    % \item (done!) Results figure, add sigma8
    % \item (more or less done!) possibly move the eigen decomposition and heat into appendix + add discussion (Nathanael I am not sure about the heat diffusion example...)
    \item Should the heat example go to appendix
    \item TK read the heat section again

\end{itemize}

@@

Before doing any numerical analysis on the sphere, one first has to choose a tessellation, i.e., an exhaustive partition of the sphere into finite area elements, where the data under study is quantized.
While our method is applicable to any pixelization of the sphere, two details depend on the chosen sampling: (i) the choice of neighbors in the construction of the graph, and (ii) the choice of parent nodes when coarsening the graph.
As HEALPix~\citep{gorski2005healpix} is the most popular sampling used in cosmology, our target application, we'll tailor the method to that particular sampling in the subsequent exposition.
%We'll note in later sections where the method should be adapted for a different sampling.

@@

HEALPix is an acronym for Hierarchical Equal Area isoLatitude Pixelization of a sphere. As suggested in the name, this pixelization produces a subdivision of a spherical surface in which each pixel covers the same surface area as every other pixel.
The lowest possible resolution is given by the base partitioning of the surface into $N_{pix} = 12$ equal sized pixels (right-most sphere in \figref{pooling}). As each pixel is subdivided in four, the second coarser resolution is $N_{pix} = N_{side}^2 \cdot 12 = 2^2 \cdot 12 = 48$ pixels (middle sphere in \figref{pooling}), the third is $N_{pix} = N_{side}^2 \cdot 12 = 4^2 \cdot 12 = 192$ pixels, etc., where $N_{side} = 1, 2, 4, 8, \ldots$ is the grid resolution parameter.
High-resolutions HEALPix maps easily reach millions of pixels. \figref{example_maps} shows two examples.
Another property of the HEALPix grid is that the pixel centers, represented by the black dots in \figref{pooling}, occur on a discrete number of rings of constant latitude (the number of which depends on the resolution).
\nati{The figure is beautiful, but should we really show it? If we target cosmologists, we can simply list some type of data. If we target a broader audience, maybe it makes sense to provide some illustration.}

@@

\label{sec:spherical_harmonics}
\begin{figure*}[!htb]
\centering
\includegraphics[width=0.95\textwidth]{figures/eigenvectors.pdf}
\caption{The first 16 Fourier modes (eigenvectors of the graph Laplacian) of a graph constructed from the Healpix sampling of the sphere. Eigenvectors 1--3 could be associated with spherical harmonics of degree $\ell=1$, eigenvectors 4--8 with degree $\ell=2$, and eigenvectors 9--15 with degree $\ell=3$. Nevertheless, the graph eigenvectors are only approximating the real spherical harmonics.}
\label{fig:graph_harmonics}
\end{figure*}

>>>>

    \item (TK) update the README for data zip - file from NP
    \item (NP) Finalise readme on Github code
    \item (NP) Finalize the data on zenodo
    \item (NP) Do the download script
    \item TK check part sphere notebook and whole sphere
    \item branch with notebooks with plots (results) NP
    \item Michael section 4.5 (efficient convolutions) and 3.2 (conv on graphs)
    \item NP move the repo to HealpixNet

    \item (TK) make the bibliography display only a few name
    % \item (yes) Michael, do you agree with the title HealpixNet? I (Michaël) extensively modified the abstract to reflect that it is a special case of graph CNNs tailored to Healpix. Hence I now agree with the title. Feel free to edit. Ideally I'd like to have "graph" in the title (graph CNNs are quite hot these days), but I couldn't find how.
    % \item Emphasize the novelty of using it on Healpix in abstract, into and rest of the paper
    \item (TK) add info on comparison with histograms, so that it's just a demonstration
    % \item Finish section on Convolutions on graphs (Michael - Nathanael)
    % \item finish section on Efficient convolutions (Michael)
    % \item (done) Plot of the new filters from Mass maps
    % \item (done!, plot to be added) check the power spectrum after smoothing
    % \item (done!) upgrade the plot of pooling (Nathanael)
    \item (TK) possibly improve plot of pooling
	% \item (done!) new plot for network structure (teaser plot) (Michael) What do you think? Todo: link the points on the sphere to the fully connected \nati{I like the plot, but I am thinking about putting multiple spheres with different maps for the convolutiion instead…}
    \item (TK NP) if we have time make a new candidate for network plot (optional)
    % \item (done!) Figure 1 to be round
    \item (first version done) conclusions
    % \item (done!) new simulations with 3 arcmin smoothing
    % \item (done!) simulation with cross-validation for the histogram
    % \item (done!) Update result section according to the new simulations...
    % \item (done!) simulation PSD
    % \item (done!) Results figure, add sigma8
    % \item (more or less done!) possibly move the eigen decomposition and heat into appendix + add discussion (Nathanael I am not sure about the heat diffusion example...)
    \item Should the heat example go to appendix
    \item TK read the heat section again
    \item TK read the result section
    \item TK check the appendix
    \item TK optional - example of data (CMB (mask) + mass map + radio intensity)
    \item TK read everything
\end{itemize}

@@

Before doing any numerical analysis on the sphere, one first has to choose a tessellation, i.e., an exhaustive partition of the sphere into finite area elements, where the data under study is quantized.
The simplicity of the spherical form belies the intricacy of global analysis on the sphere. There is no known point set which achieves the analogue of uniform sampling in Euclidean space and allows exact and invertible discrete spherical harmonic decompositions of arbitrary but band-limited functions. Any existing proposition of practical schemes for the discrete treatment of such functions on the sphere introduces some systematic error dependent on the global properties of the point set.
While our method is applicable to any pixelization of the sphere, two details depend on the chosen sampling: (i) the choice of neighbors in the construction of the graph, and (ii) the choice of parent nodes when coarsening the graph.
As HEALPix~\citep{gorski2005healpix} is the most popular sampling used in cosmology, our target application, we'll tailor the method to that particular sampling in the subsequent exposition.
%We'll note in later sections where the method should be adapted for a different sampling.

@@

HEALPix is an acronym for Hierarchical Equal Area isoLatitude Pixelization of a sphere. As suggested in the name, this pixelization produces a hierarchical subdivision of a spherical surface in which each pixel covers the same surface area as every other pixel.
A hierarchy is desired for data locality in computer memory. Equal area is advantageous because white noise generated by the signal receiver gets integrated exactly into white noise in the pixel space. Isolatitude is essential for the implementation of a fast spherical transform. HEALPix is the sole pixelization scheme which satisfies those three properties.

The lowest possible resolution is given by the base partitioning of the surface into $N_{pix} = 12$ equal sized pixels (right-most sphere in \figref{pooling}). As each pixel is subdivided in four, the second coarser resolution is $N_{pix} = N_{side}^2 \cdot 12 = 2^2 \cdot 12 = 48$ pixels (middle sphere in \figref{pooling}), the third is $N_{pix} = N_{side}^2 \cdot 12 = 4^2 \cdot 12 = 192$ pixels, etc., where $N_{side} = 1, 2, 4, 8, \ldots$ is the grid resolution parameter.
High-resolutions HEALPix maps easily reach millions of pixels. \figref{example_maps} shows two examples.
\nati{The figure is beautiful, but should we really show it? If we target cosmologists, we can simply list some type of data. If we target a broader audience, maybe it makes sense to provide some illustration.}

@@

\label{sec:spherical_harmonics}
\begin{figure*}[!htb]
\centering
\includegraphics[width=0.95\textwidth]{figures/eigenvectors.pdf}
\caption{The first 16 Fourier modes (eigenvectors of the graph Laplacian) of a graph constructed from the Healpix sampling of the sphere. Eigenvectors 1--3 could be associated with spherical harmonics of degree $\ell=1$, eigenvectors 4--8 with degree $\ell=2$, and eigenvectors 9--15 with degree $\ell=3$. Nevertheless, the graph eigenvectors are only approximating the real spherical harmonics. \todo{show group of eigenvectors as indication of mode $\ell$. Either show eigenvalues on this plot or add a plot of the eigenvalues themselves.}}
\label{fig:graph_harmonics}
\end{figure*}
